---
title: "Assignment 1"
author: "Mary Geschwindt"
date: "9/5/2020"
output: html_document
---

# Importing the TidyCensus and Tidyverse Libraries

```{r}
library(tidycensus)
library(tidyverse)
```

# Creating a Dataset

## Area of Interest

In considering which variables to select for my data set for this assignment, I found it useful to first come up with a topic I am interested in studying. The topic that I have considered for assignment 1 is:

**The relationship between age and travel distance to work for women who walked to work in New York State in 2018.**

## Variables

Based on my research question, I know that I need to use the following person-level variables:

1. Sex (categorical): SEX
2. Means of transportation to work (categorical): JWTR
3. Age (continuous): AGEP
4. Travel time to work (continuous): JWMNP

I will additionally include the following variable which may be interesting for further study:

1. Income (continuous): PINCP

## Defining the Variables

Now that I've selected which variables to use, I'll load the variables to create the data frame.

```{r}
women_walking2018 <- get_pums(variables = c("SEX",
                                      "JWTR",
                                      "JWMNP",
                                      "AGEP",
                                      "PINCP"),
                        state = "NY",
                        year = 2018,
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(SEX == 2,
         JWTR == 10,
         JWMNP > 0,
         AGEP > 18,
         PINCP > 0) %>%
  select(SEX_label, JWTR_label, JWMNP, AGEP, PINCP)
```

# Results

## Data Frame Display

The following display shows the first few rows of the data frame that I created based on the variables loaded:

```{r}
women_walking2018
```

## Observations

There are 2,553 rows, or observations in this data frame. The observations are taken at the person-level, meaning there are 2,553 people included in this analysis.

# Saving CVS File

Before I commit and push this R Marked Down file to Github, I'll save the data loaded as a CSV file:

```{r}
write.csv(women_walking2018, "women_walking2018.csv")
```

