---
title: "An Analysis of Women's Commutes in New York City"
author: "Mary Geschwindt"
date: "10/4/2020"
output:
  html_document:
    theme: paper
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Document Setup

```{r load libraries, message=FALSE, results='hide'}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(scales)
library(knitr)
library(jtools)
library(interactions)
```

```{r load data, message=FALSE}
women_data <- read.csv("women_nyctransport2018.csv")

data.frame(women_data)
```

```{r recode, message=FALSE, results='hide'}
women_data <- read_csv("women_nyctransport2018.csv") %>%
  mutate(JWTR_label = case_when(
    JWTR_label=="Bicycle"~"Bicycle",
    JWTR_label=="Bus or trolley bus"~"Bus",
    JWTR_label=="Car, truck, or van"~"1Car",
    JWTR_label=="Ferryboat"~"Ferry",
    JWTR_label=="Other method"~"Other",
    JWTR_label=="Railroad"~"Train",
    JWTR_label=="Subway or elevated"~"Subway",
    JWTR_label=="Taxicab"~"Taxi",
    JWTR_label=="Walked"~"Walked")) %>%
  mutate(DPHY_label = case_when(
    DPHY_label=="Yes"~"Physically Disabled",
    DPHY_label=="No"~"Not Phsycailly Disabled"))
```


## Research Question and Hypothesis

For this project, I was curious about women's commutes in New York City. I wanted to study what types of modes of transportation are women most likely to use, and whether having a physical disability impacts access to certain transportation modes. My research question is what is the impact on a woman's travel time to work in New York City based on her mode of transportation and disability status?

My hypothesis is that women who live and commute to work in New York City may have varied travel times to work depending on their means of transportation, their age, their income, and whether or not they have a physical disability. I was curious about race as a variable for several of the steps in the study, so I included that variable in my dataset as well. I hypothesize that these variables may have positive or negative impacts on women's lives depending on the lengths of their commutes.

This research question is important because it is a means of understanding how equitable the transportation infrastructures are in New York City. There are many ways to define what equitable means; one can look at equity in terms of gender, race, age, physical ability, income, and other variables. This assignment cannot, and does not seek to, determine whether or not New York's transportation infrastructures serve all people in the City equally. But, it does look at one specific population to try to determine whether some of the variables do make a difference in some women having an unequal experience.

## Dataset

### Population

The population is women in who live in New York City who have a commute to work. The data is taken at the person-level, and there are 327 observations in the dataset. This means that there are 327 women who are included in this analysis. All New York City Public Use Microdata Areas (PUMAs) were used to define the geographic boundaries for the sample of women counted.

### Variables

Using the US Census 1-year American Community Survey from 2018 as my data source, I selected the following variables to test my hypothesis:

**Means of transportation to work (categorical): JWTR**
I selected means of transportation to work because this is a variable which may impact commute times, depending on which mode of transportation is used. This variable can also be impacted by other variables, such as income, which may influence which means of transportation a woman in this sample is able to afford to take daily.

**Age (continuous): AGEP**
Age was selected because a woman's age may be related to her income and therefore what means of transportation she can afford. It may also be related to her physical abilities in taking certain kinds of transportation.

**Travel time to work (continuous): JWMNP**
Travel time to work was selected because there may be relationship between all other variables and how fast a woman can get to work given the other selected variables. I selected travel time under the assumption that it could be one of the indicators for how equitable a transportation system is.

**Ambulatory difficulty (categorical): DPHY**
Ambulatory difficulty, or physical disability, was selected because it may be related to a woman's inability to take certain types of transportation modes, which in turn could impact travel times.

**Race (categorical): RAC1P**
Race was selected because I was curious if there was a relationship between access to a mode of transportation and race based on a woman's physical location - the reason for selecting this variable presents a slightly different hypothesis, which is that race may be related to where a  woman lives, and therefore determines what transportation infrastructures she may have access to in her home neighborhood. Where a woman lives and what she has access to might be related to her travel time to work. This variable is interesting to look at in terms of correlation, but it requires further study for a more specific and slightly different hypothesis related to how equitable transportation infrastructures are for women in New York City.

**Income (continuous): PINCP**
Income was selected because it may be related to what means of transportation a women might be able to afford in New York City; for example, the costs associated with owning and maintaining a car to drive to work in NYC are much higher than owning a bicycle which a woman can ride to work instead.

## Methods

### Distributions & Proportions

Once the population and variables were determined and the dataset created, my first method was to understand what the distributions and proportions of this sample looked like. To do this, I took a different approach depending on whether the variables were continuous or categorical.

For continuous variables, I calculated the summary statistics, including the sample mean, sample standard deviation, the 95-percent confidence interval for the population mean, and the interquartile range. The results for the continuous variables, which include age, income, and travel time to work, are as follows:

```{r summary statistics calculation, message=FALSE}
age_summary <- summary(women_data$AGEP)
age_sd <- sd(AGEP, na.rm=TRUE)
```


## Results



## Discussion & Limitations

### Transportation Costs

Transportation costs is a variable which I wish I was able to include from the beginning, but which was not available as a variable in the 1-year ACS from 2018. I think that looking at the transportation costs would strengthen the question of whether or not a system is equitable by comparing transportation costs with income.

### Race Variable

As previously mentioned, I had included race as a variable early on in creating my dataset. I initially wanted to understand if there was a relationship between race, a woman's means of transportation to work, and travel time to work. In pursuing the question of equity in NYC transportation infrastructures, I assumed race would be a necessary variable to include.

After completing the various tests and interpreting the results, I came to the conclusion that I need to include more variables when looking at race in order to answer a slightly different question. Looking at race on its own with respect to these other variables is not enough to reach a conclusion about equity. To address this additional question, altering this study to include a variable for neighborhood would be necessary in determining if there is even a correlation between race, where a woman lives, her access to transportation modes, and travel time to work.

### Disability Variables

For the sake of simplicity for the first assignment, I had only included ambulatory difficulty when looking at disability. In order to have a more well-rounded approach in determining whether or not a transportation system is equitable for disabled users, it is essential to include other disability variables, such as for vision, hearing, and cognitive disabilities. By including a wider range of disabilities, the conclusion of whether or not a transportation infrastructural system is equitable would be more in line with including everyone. It is not possible to talk about equity for all New Yorkers if entire groups of people are excluded.

### Sample Size

There are only 327 observations in my dataset. Had the sample size been larger, it may have been possible for other relationships to emerge. For example, there were only XX observations in my sample who have a physical disability, compared to the remaining XXX observations who do not have a physical disability. Had the sample size been greater, I suspect there would have been more women who had a phsyical disability.

### Gender Variable

In order to more accurately address the question "Is it equitable?", my analysis may benefit from making a comparison between women and men, with men being the reference value for a gender variable. The question I studied for this project was mainly concerned with women and whether or not they had a disability, but determining true equity requires including many other variables related to intersectional characteristics. 

## Conclusion

In New York City, about a quarter of the 472 subway stations are ADA accessible as of this year (an infographic depicting which stations are ADA accessible can be seen here: https://gothamist.com/news/infographic-how-much-nyc-subway-accessible). Poorly maintained bike lanes cause bumpy and uncomfortable rides for abled persons, yet they nearly eliminate the option of biking for some disabled people. Walking is a mode of transportation employed by many of the City's residents, but not all sidewalks are well-maintained, and there are instances of sidewalk obstructions which make it impossible for a person in a wheelchair or someone with a stroller to pass.

All of this is to say that New York City is comprised of a diverse population with many needs. It is the responsibility of the City to provide for the needs of everyday New Yorkers in a way that includes providing equal transportation service to everyone, regardless of gender, ability, age, race, socioeconomic status, and so on. It is necessary to look at this kind of data to determine whether or not the City's transportation infrastructure is equitable. Unequal service for people who live in the same city and are served by the same government agencies and officials is unequal and unjust.

People should care about this analysis and other similar analyses because the role of public transportation and transportation infrastructures maintained by government agencies (such as sidewalk and roadway conditions) is to serve the public. If there are members of the public which have a more difficult time using certain types of infrastructures, then 